---
title: "Group Project Advanced Statistical Programming using R"
author: "Tiago HAGEN, Marco ANTON, Nizami AZIZOV and Louisa DJEBROUN"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    code-tools: true
    number-sections: true
    theme: cosmo
execute:
  warning: false
  message: false
editor: 
  markdown: 
    wrap: 72
---

```{r setup}
#| include: false
library(tidyverse)
library(scales)
library(knitr)
library(patchwork)

```

```{r load-data}
#| include: false
df_raw <- read_csv("Data/Social_Media_Impact_on_Daily_Lives.csv")

# Rename columns to short
df <- df_raw %>% 
  rename(
    idx             = `...1`,
    name            = `Enter your name`,
    age             = `Enter your age (Ex: 20)`,
    gender          = `What is your gender`,
    profession      = `What is your profession?`,
    prof_hours      = `On an average how many hours a day you spend on your profession?`,
    platforms       = `What social networking sites do you use?`,
    social_hours    = `On an average how many hours a day do you spend on these sites?`,
    family_hours    = `On an average how many hours a day do you spend with your family?`,
    why_use         = `Why do you use these social networking sites?`,
    benefits        = `What are personal benefits of  using social networking sites?`,
    when_access     = `When do you access social media websites?`,
    privacy_effect  = `Do you think privacy policies are effective in social networking sites?`,
    ad_influence    = `Does social media advertising influence the purchase of your product?`,
    forget_time     = `When you use social media did you forget about time?`,
    addicted        = `Do you consider yourself addicted to social media?`
  ) %>% 
  # Fix the one row where profession = "male" 
  mutate(profession = if_else(profession == "male", NA_character_, profession))

# convert to numeric midpoints for regression
hour_midpoint <- function(x) {
  case_when(
    x == "0-2"  ~ 1,
    x == "1-2"  ~ 1.5,
    x == "2-4"  ~ 3,
    x == "4-5"  ~ 4.5,
    x == "4-6"  ~ 5,
    x == "5-6"  ~ 5.5,
    x == "6-8"  ~ 7,
    x == "6+"   ~ 7,
    x == "8+"   ~ 9,
    TRUE        ~ NA_real_
  )
}

df <- df %>% 
  mutate(
    social_hours_n = hour_midpoint(social_hours),
    family_hours_n = hour_midpoint(family_hours),
    addicted_bin   = case_when(
      addicted == "yes"      ~ 1,
      addicted == "no"       ~ 0,
      addicted == "not sure" ~ NA_integer_
    )
  )
```

# Dataset & research questions

## Reliability of License , strong source

We searched for a more complex dataset so we could analyse the
relationships between columns and not only one column to the others.

##Lack of documentation There is no official documentation for the
spreadsheet. They provided the data just following the study. The
problems with the data have already been stated.

## Source and licence\_

The dataset is titled "Social Media Impact on Daily Lives" and can be
downloaded
[here](https://data.mendeley.com/datasets/9s2nmbxkrs/2/files/6647758b-2eca-4158-a49e-a82611be62f0).
It was collected via a structured online survey on January 2025.

Licence: CC BY 4.0

## What is in it

The dataset contains `r nrow(df)` rows and `r ncol(df_raw)` columns.
Each row represents one survey respondent. The key variables are:

| Variable | Rename | Type | Description |
|------------------|------------------|------------------|------------------|
| `Enter your age (Ex: 20)` | age | numeric | Respondent's age in years |
| `What is your gender` | gender | categorical | Male / Female |
| `What is your profession?` | profession | categorical | Student / Employee |
| `On an average how many hours a day you spend on these sites?` | social_hours | ordinal | Daily hours spent on social media |
| `On an average how many hours a day do you spend with your family?` | family_hours | ordinal | Daily hours spent with family |
| `Does social media advertising influence the purchase of your product?` | ad_influence | ordinal | Whether social media advertising influences purchases |
| `When you use social media did you forget about time?` | forget_time | categorical | Whether the user forgets time while on social media |
| `Do you consider yourself addicted to social media?` | addicted | categorical | Self-reported addiction: Yes / No / Not sure |

## Research questions

### Descriptive & Exploratory

**Do younger people spend more time on social media?**\
We aim to describe the distribution of daily social media usage across
age groups and report whether a negative or positive association exists
between age and screen time in this sample.

### Explanatory & Exploratory

**Would people with higher screen time consider themselves addicted?**\
We will examine people who spend more hours on social media are more
likely to report feeling addicted.

### Explanatory & Confirmatory

**Do people with higher social media screen time spend less time with
their family?**\
We will examine whether daily social media usage and daily family time
are negatively correlated, controlling informally for profession.

### Explanatory & Exploratory

**Does the perceived influence of social media advertising on purchasing
decisions differ between men and women?**\
We will compare the distribution of the `ad_influence` variable across
genders and assess whether the difference is substantively meaningful.

### Predictive & Exploratory

**Can we predict that someone considers themselves addicted to social
media?**\
Using logistic regression, we will model the binary outcome `addicted`
as a function of age, gender, social media hours, and perceived
benefits.

## Focus questions:

**Can we predict that someone considers themselves addicted to social
media?** and **Do people with higher social media screen time spend less
time with their family?**\

## Context and target audience

This analysis explores how social media usage relates to: - Addiction
perception - Family/social life balance - Demographic differences (age,
gender) - Consumer behavior (advertising influence)

Target audience: Product managers, Social Media managers, policy makers
(political and social)\
Interactive Dashboard would be suitable for this audience.
