---
title: "Group Project Advanced Statistical Programming using R"
author: "Tiago HAGEN, Marco ANTON, Nizami AZIZOV and Louisa DJEBROUN"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    code-tools: true
    number-sections: true
    theme: cosmo
execute:
  warning: false
  message: false
editor: 
  markdown: 
    wrap: 72
---

# Initial data analysis

```{r glimpse}
library(tidyverse)
library(plotly)
glimpse(df)
```

```{r summary-stats}
df %>% 
  select(age, social_hours_n, family_hours_n) %>% 
  summary()
```

```{r categorical-counts}
#| label: tbl-counts
#| tbl-cap: "Response counts for key categorical variables"

bind_rows(
  df %>%  count(Variable = "Gender",       Value = gender)      %>%  rename(Count = n),
  df %>%   count(Variable = "Profession",   Value = profession)  %>%  rename(Count = n),
  df %>%  count(Variable = "Addicted",     Value = addicted)    %>%  rename(Count = n),
  df %>%  count(Variable = "Forget time",  Value = forget_time) %>%  rename(Count = n)
) %>% 
  drop_na(Value) %>% 
  kable()
```

```{r}
df_raw %>% 
  summarise(across(everything(), ~ sum(is.na(.)))) %>% 
  pivot_longer(everything(), names_to = "Variable", values_to = "Missing") %>% 
  filter(Missing > 0)


df_raw %>% 
  filter(`What is your profession?` == "male")
```

**Missing values:** The `name` column has 1 missing entry, and 1 row had
`profession = "male"` (a clear data-entry error, recoded to `NA`). All
other key variables are complete (0 missing values). Hour variables are
encoded as intervals (e.g. `"3-4"` hours) rather than continuous values,
which limits precision.

## Distribution plot

```{r fig-distribution}
#| label: fig-dist
#| fig-cap: "Distribution of daily social media usage and age by self-reported addiction status"
#| fig-width: 10
#| fig-height: 4

p1 <- df %>% 
  filter(!is.na(addicted)) %>% 
  mutate(
    social_hours = factor(social_hours, levels = c("1-2","3-4","4-5","5-6","6+"))
  ) %>% 
  count(social_hours, addicted) %>% 
  group_by(social_hours) %>% 
  mutate(pct = n / sum(n)) %>% 
  ggplot(aes(x = social_hours, y = pct, fill = addicted)) +
  geom_col(position = "fill") +
  scale_y_continuous(labels = percent_format()) +
  scale_fill_manual(values = c("yes" = "#E07070", "no" = "#70B8E0", "not sure" = "#C0C0C0")) +
  labs(
    title = "Addiction by Daily Screen Time",
    x = "Hours/day on social media",
    y = "Proportion",
    fill = "Addicted?"
  ) +
  theme_minimal()

p2 <- df %>% 
  ggplot(aes(x = age)) +
  geom_histogram(binwidth = 2, fill = "#4C9BE8", colour = "white") +
  labs(title = "Age Distribution of Respondents", x = "Age (years)", y = "Count") +
  theme_minimal()

p2 + p1
```

```{r}
#| fig-width: 14
#| fig-height: 16
p1 <- ggplot(df, aes(
    x = social_hours,
    fill = addicted
  )) +
  geom_bar(position = "dodge") +
  theme_minimal() +
  labs(
    title = "Relationship: Social Media Usage vs. Self-Reported Addiction",
    x = "Average Daily Hours Spent on Social Media",
    y = "Number of Respondents",
    fill = "Considered Addicted?"
  )

p2 <- df %>% 
  filter(!is.na(social_hours), !is.na(family_hours)) %>% 
  mutate(
    social_hours = factor(social_hours, levels = c("1-2", "3-4", "4-5", "5-6", "6+")),
    family_hours = factor(family_hours, levels = c("1-2", "3-4", "4-5", "5-6", "6+"))
  ) %>% 
  ggplot(aes(x = social_hours, fill = family_hours)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(
    title = "Social Media Hours vs. Family Time",
    x = "Daily social media hours",
    y = "Proportion",
    fill = "Family hours/day"
  ) +
  theme_minimal()

ggplotly(p2)


p3 <- df %>%
  count(
    gender = gender,
    advertising_influence = ad_influence
  ) %>%
  group_by(gender) %>%
  mutate(percent = n / sum(n) * 100) %>%
  ggplot(aes(
    x = advertising_influence,
    y = percent,
    fill = gender
  )) +
  geom_col(position = "dodge") +
  labs(
    title = "Influence of Social Media Advertising by Gender",
    x = "Advertising influence on purchase decision",
    y = "Percentage within gender group",
    fill = "Gender"
  ) +
  theme_minimal()

p4 <- df %>% 
  mutate(
    age_group = cut(age, breaks = c(13, 20, 25, 30, 50),
                    labels = c("13-20", "21-25", "26-30", "30+")),
    social_hours = factor(social_hours, levels = c("1-2", "3-4", "4-5", "5-6", "6+"))
  ) %>% 
  ggplot(aes(x = age_group, fill = social_hours)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = percent_format()) +
  labs(
    title = "Age vs Daily Social Media Usage",
    x = "Age group",
    y = "Proportion",
    fill = "Hours/day"
  ) +
  theme_minimal()

p5 <- df %>% 
  filter(!is.na(addicted_bin)) %>% 
  ggplot(aes(x = social_hours_n, fill = factor(addicted_bin))) +
  geom_histogram(position = "fill", binwidth = 1) +
  scale_y_continuous(labels = percent_format()) +
  scale_fill_manual(values = c("0" = "#70B8E0", "1" = "#E07070"),
                    labels = c("0" = "No", "1" = "Yes")) +
  labs(
    title = "Proportion Addicted by Screen Time",
    x = "Hours/day (midpoint)",
    y = "Proportion",
    fill = "Addicted?"
  ) +
  theme_minimal()


(p4 + p1) / (p2 + p3) / (p5)

p6
```

## Data quality

- Intervals, not continuous values: all time variables (`social_hours`,
  `family_hours`, `prof_hours`) are in the form of strings (for example,
  `“3-4”`).
- 78% of respondents are students, which limits generalisability to
  employed populations.
- One row has `profession = "male"` recoded to `NA`.
- The response options are Yes / No / Not sure. The "Not sure" group
  (`n = 125`, 24%) is substantial and will be excluded from binary
  logistic regression, reducing the effective sample.
