---
title: "Group Project Advanced Statistical Programming using R"
author: "Tiago HAGEN, Marco ANTON, Nizami AZIZOV and Louisa DJEBROUN"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    code-tools: true
    number-sections: true
    theme: cosmo
execute:
  warning: false
  message: false
editor: 
  markdown: 
    wrap: 72
---
```{r setup}
#| include: false
source("setup.R")
```
# Initial data analysis

```{r glimpse}
glimpse(df)
```

```{r summary-stats}
df %>% 
  select(age, social_hours_n, family_hours_n) %>% 
  summary()
```

```{r categorical-counts}
#| label: tbl-counts
#| tbl-cap: "Response counts for key categorical variables"

bind_rows(
  df %>%  count(Variable = "Gender",       Value = gender)      %>%  rename(Count = n),
  df %>%   count(Variable = "Profession",   Value = profession)  %>%  rename(Count = n),
  df %>%  count(Variable = "Addicted",     Value = addicted)    %>%  rename(Count = n),
  df %>%  count(Variable = "Forget time",  Value = forget_time) %>%  rename(Count = n)
) %>% 
  drop_na(Value) %>% 
  kable()
```

```{r}
df_raw %>% 
  summarise(across(everything(), ~ sum(is.na(.)))) %>% 
  pivot_longer(everything(), names_to = "Variable", values_to = "Missing") %>% 
  filter(Missing > 0)

df_raw %>% 
  filter(`What is your profession?` == "male")
```

**Missing values:** The `name` column has 1 missing entry, and 1 row had
`profession = "male"` (a clear data-entry error, recoded to `NA`). All
other key variables are complete (0 missing values). Hour variables are
encoded as intervals (e.g. `"3-4"` hours) rather than continuous values,
which limits precision.

## Distribution plot

```{r fig-distribution}
#| label: fig-dist
#| fig-cap: "Distribution of daily social media usage and age by self-reported addiction status"
#| fig-width: 10
#| fig-height: 4

p1 <- df %>% 
  filter(!is.na(addicted)) %>% 
  mutate(
    social_hours = factor(social_hours, levels = c("1-2","3-4","4-5","5-6","6+"))
  ) %>% 
  count(social_hours, addicted) %>% 
  group_by(social_hours) %>% 
  mutate(pct = n / sum(n)) %>% 
  ggplot(aes(x = social_hours, y = pct, fill = addicted)) +
  geom_col(position = "fill") +
  scale_y_continuous(labels = percent_format()) +
  scale_fill_manual(values = c("yes" = "#E07070", "no" = "#70B8E0", "not sure" = "#C0C0C0")) +
  labs(
    title = "Addiction by Daily Screen Time",
    x = "Hours/day on social media",
    y = "Proportion",
    fill = "Addicted?"
  ) +
  theme_minimal()

p2 <- df %>% 
  ggplot(aes(x = age)) +
  geom_histogram(binwidth = 2, fill = "#4C9BE8", colour = "white") +
  labs(title = "Age Distribution of Respondents", x = "Age (years)", y = "Count") +
  theme_minimal()

p2 + p1
```
**Interpretation:** The left histogram reveals that the respondents' ages are heavily concentrated between 18 and 26 years, reflecting a predominantly student-based demographic. The right proportion bar chart demonstrates a clear dependency trend: while users spending 1–3 hours daily mostly report not feeling addicted, the proportion of self-reported addiction ("Yes") expands rapidly as daily screen time exceeds 4 hours.

```{r}
#| label: fig-ad-influence
#| fig-cap: "Social media advertising shows a similar distribution of influence on purchasing decisions for both male and female respondents."
#| fig-alt: "A grouped bar chart comparing the influence of social media advertising on purchasing decisions between men and women. The distributions are very similar across all categories."
df %>%
  count(gender, ad_influence) %>%
  group_by(gender) %>%
  mutate(percent = n / sum(n) * 100) %>%
  ggplot(aes(x = ad_influence, y = percent, fill = gender)) +
  geom_col(position = "dodge") +
  # Colorblind-safe color palette
  scale_fill_viridis_d(option = "cividis", end = 0.8) +
  theme_minimal() +
  labs(
    #  Title states the finding, not just the variables
    title = "Advertising influences purchasing decisions similarly across genders",
    subtitle = "Self-reported influence levels by gender",
    # Dropped redundant x-axis label (y = NULL logic applied to x)
    x = NULL, 
    #  Axis label with units (%)
    y = "Proportion within gender group (%)",
    fill = "Gender",
    #  Data source caption
    caption = "Data Source: Social Media Impact on Daily Lives (2025)"
    )
```
**Interpretation:** This grouped bar chart compares the self-reported influence of social media advertising on purchasing behavior across genders. The distributions are remarkably similar, with both males and females most frequently reporting an "often" or "sometimes" influence on their consumer choices. This suggests that gender alone does not explain differences in susceptibility to digital marketing within this sample.

```{r}
#| label: fig-social-addiction
#| fig-width: 10
#| fig-height: 6
p1 <- ggplot(df, aes(
    x = social_hours,
    fill = addicted
  )) +
  geom_bar(position = "dodge") +
  theme_minimal() +
  labs(
    title = "Relationship: Social Media Usage vs. Self-Reported Addiction",
    x = "Average Daily Hours Spent on Social Media",
    y = "Number of Respondents",
    fill = "Considered Addicted?"
  )
p1
```
**Interpretation:** By viewing absolute counts rather than percentages, this chart highlights that the bulk of our sample falls into the 2–5 daily hours range. Within these moderate-to-high usage brackets, the number of users who acknowledge addiction surpasses those who deny it. Notably, the "Not sure" group remains present across all tiers, underscoring the subjective difficulty many individuals face when assessing their own digital dependency.

```{r}
#| label: fig-family-time
#| fig-width: 10
#| fig-height: 6
p2_interaktiv <- df %>% 
  filter(!is.na(social_hours), !is.na(family_hours)) %>% 
  mutate(
    social_hours = factor(social_hours, levels = c("1-2", "3-4", "4-5", "5-6", "6+")),
    family_hours = factor(family_hours, levels = c("1-2", "3-4", "4-5", "5-6", "6+"))
  ) %>% 
  count(social_hours, family_hours) %>% 
  group_by(social_hours) %>% 
  mutate(percent = n / sum(n) * 100) %>% 
  ggplot(aes(x = social_hours, y = percent, fill = family_hours, 
             text = paste0("Family Time: ", family_hours, "<br>",
                           "Anteil: ", round(percent, 1), " %"))) +
  geom_col() +
  labs(
    title = "Social Media Hours vs. Family Time",
    x = "Daily social media hours",
    y = "Proportion (%)",
    fill = "Family hours/day"
  ) +
  theme_minimal()

ggplotly(p2_interaktiv, tooltip = "text")
```
**Interpretation:** This interactive proportion chart examines whether high screen time displaces offline family interaction. Contrary to the assumption of a strict negative correlation, the distribution of family time remains surprisingly stable across different social media usage brackets. This indicates that screen time alone does not directly cannibalize family hours, suggesting that factors such as academic workload or living arrangements may act as stronger confounders.

```{r}
#| label: fig-age-usage
#| fig-width: 10
#| fig-height: 6
p4 <- df %>% 
  mutate(
    age_group = cut(age, breaks = c(13, 20, 25, 30, 50),
                    labels = c("13-20", "21-25", "26-30", "30+")),
    social_hours = factor(social_hours, levels = c("1-2", "3-4", "4-5", "5-6", "6+"))
  ) %>% 
  ggplot(aes(x = age_group, fill = social_hours)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = percent_format()) +
  labs(
    title = "Age vs Daily Social Media Usage",
    x = "Age group",
    y = "Proportion",
    fill = "Hours/day"
  ) +
  theme_minimal()
p4
```
**Interpretation:** This visualization categorizes respondents into age brackets to assess generational differences in screen time. Interestingly, daily usage distributions are relatively consistent across all age groups, including older respondents (30+). This finding challenges the stereotype that heavy social media consumption is strictly confined to teenagers and young adults, showing widespread digital engagement across the entire sample.

```{r}
#| label: fig-binary-addiction
#| fig-width: 10
#| fig-height: 6
p5 <- df %>% 
  filter(!is.na(addicted_bin)) %>% 
  ggplot(aes(x = social_hours_n, fill = factor(addicted_bin))) +
  geom_histogram(position = "fill", binwidth = 1) +
  scale_y_continuous(labels = percent_format()) +
  scale_fill_manual(values = c("0" = "#70B8E0", "1" = "#E07070"),
                    labels = c("0" = "No", "1" = "Yes")) +
  labs(
    title = "Proportion Addicted by Screen Time",
    x = "Hours/day (midpoint)",
    y = "Proportion",
    fill = "Addicted?"
  ) +
  theme_minimal()
p5
```
**Interpretation:** By isolating the binary outcome (excluding "Not sure" responses and using numeric hour midpoints), this plot clearly illustrates the escalating probability of self-perceived addiction. At approximately 4 to 5 hours of daily use, the sample crosses a critical tipping point where more than 50% of users consider themselves addicted. Beyond 6 hours daily, self-reported addiction becomes the overwhelming majority, visually validating our predictive modeling hypothesis.

## Data quality
- Intervals, not continuous values: all time variables (`social_hours`,
  `family_hours`, `prof_hours`) are in the form of strings (for example,
  `“3-4”`).
- 78% of respondents are students, which limits generalisability to
  employed populations
- One row has `profession = "male"` recoded to `NA`.
- The response options are Yes / No / Not sure. The "Not sure" group
  (`n = 125`, 24%) is substantial and will be excluded from binary
  logistic regression, reducing the effective sample.
