---
title: "Group Project Advanced Statistical Programming using R"
author: "Tiago HAGEN, Marco ANTON, Nizami AZIZOV and Louisa DJEBROUN"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    code-tools: true
    number-sections: true
    theme: cosmo
execute:
  warning: false
  message: false
editor: 
  markdown: 
    wrap: 72
---

```{r setup}
#| include: false
library(tidyverse)
library(scales)
library(knitr)
library(patchwork)

```

```{r load-data}
#| include: false
df_raw <- read_csv("Data/Social_Media_Impact_on_Daily_Lives.csv")

# Rename columns to short
df <- df_raw %>% 
  rename(
    idx             = `...1`,
    name            = `Enter your name`,
    age             = `Enter your age (Ex: 20)`,
    gender          = `What is your gender`,
    profession      = `What is your profession?`,
    prof_hours      = `On an average how many hours a day you spend on your profession?`,
    platforms       = `What social networking sites do you use?`,
    social_hours    = `On an average how many hours a day do you spend on these sites?`,
    family_hours    = `On an average how many hours a day do you spend with your family?`,
    why_use         = `Why do you use these social networking sites?`,
    benefits        = `What are personal benefits of  using social networking sites?`,
    when_access     = `When do you access social media websites?`,
    privacy_effect  = `Do you think privacy policies are effective in social networking sites?`,
    ad_influence    = `Does social media advertising influence the purchase of your product?`,
    forget_time     = `When you use social media did you forget about time?`,
    addicted        = `Do you consider yourself addicted to social media?`
  ) %>% 
  # Fix the one row where profession = "male" 
  mutate(profession = if_else(profession == "male", NA_character_, profession))

# convert to numeric midpoints for regression
hour_midpoint <- function(x) {
  case_when(
    x == "0-2"  ~ 1,
    x == "1-2"  ~ 1.5,
    x == "2-4"  ~ 3,
    x == "4-5"  ~ 4.5,
    x == "4-6"  ~ 5,
    x == "5-6"  ~ 5.5,
    x == "6-8"  ~ 7,
    x == "6+"   ~ 7,
    x == "8+"   ~ 9,
    TRUE        ~ NA_real_
  )
}

df <- df %>% 
  mutate(
    social_hours_n = hour_midpoint(social_hours),
    family_hours_n = hour_midpoint(family_hours),
    addicted_bin   = case_when(
      addicted == "yes"      ~ 1,
      addicted == "no"       ~ 0,
      addicted == "not sure" ~ NA_integer_
    )
  )
```

# Dataset & research questions

## Source and licence

The dataset is titled "Social Media Impact on Daily Lives" and can be
downloaded
[here](https://data.mendeley.com/datasets/9s2nmbxkrs/2/files/6647758b-2eca-4158-a49e-a82611be62f0).
It was collected via a structured online survey on January 2025.

Licence: CC BY 4.0

## What is in it

The dataset contains `r nrow(df)` rows and `r ncol(df_raw)` columns.
Each row represents one survey respondent. The key variables are:

| Variable | Rename | Type | Description |
|------------------|------------------|------------------|------------------|
| `Enter your age (Ex: 20)` | age | numeric | Respondent's age in years |
| `What is your gender` | gender | categorical | Male / Female |
| `What is your profession?` | profession | categorical | Student / Employee |
| `On an average how many hours a day you spend on these sites?` | social_hours | ordinal | Daily hours spent on social media |
| `On an average how many hours a day do you spend with your family?` | family_hours | ordinal | Daily hours spent with family |
| `Does social media advertising influence the purchase of your product?` | ad_influence | ordinal | Whether social media advertising influences purchases |
| `When you use social media did you forget about time?` | forget_time | categorical | Whether the user forgets time while on social media |
| `Do you consider yourself addicted to social media?` | addicted | categorical | Self-reported addiction: Yes / No / Not sure |

## Research questions

### Descriptive

**Do younger people spend more time on social media?**\
We aim to describe the distribution of daily social media usage across
age groups and report whether a negative or positive association exists
between age and screen time in this sample.

### Explanatory

**Would people with higher screen time consider themselves addicted?**\
We will examine people who spend more hours on social media are more
likely to report feeling addicted.

### Explanatory

**Do people with higher social media screen time spend less time with their family?**\
We will examine whether daily social media usage and daily family time
are negatively correlated, controlling informally for profession.

### Explanatory

**Does the perceived influence of social media advertising on purchasing decisions differ between men and women?**\
We will compare the distribution of the `ad_influence` variable across
genders and assess whether the difference is substantively meaningful.

### Predictive

**Can we predict that someone considers themselves addicted to social media?**\
Using logistic regression, we will model the binary outcome `addicted`
as a function of age, gender, social media hours, and perceived
benefits.


## Context and target audience

This analysis explores how social media usage relates to: 
- Addiction perception 
- Family/social life balance 
- Demographic differences (age, gender) 
- Consumer behavior (advertising influence)

Target audience: Product managers, Social Media managers, policy makers (political and social) \
Interactive Dashboard would be suitable for this audience.

# Initial data analysis

```{r glimpse}
glimpse(df)
```

```{r summary-stats}
df %>% 
  select(age, social_hours_n, family_hours_n) %>% 
  summary()
```

```{r categorical-counts}
#| label: tbl-counts
#| tbl-cap: "Response counts for key categorical variables"

bind_rows(
  df %>%  count(Variable = "Gender",       Value = gender)      %>%  rename(Count = n),
  df %>%   count(Variable = "Profession",   Value = profession)  %>%  rename(Count = n),
  df %>%  count(Variable = "Addicted",     Value = addicted)    %>%  rename(Count = n),
  df %>%  count(Variable = "Forget time",  Value = forget_time) %>%  rename(Count = n)
) %>% 
  drop_na(Value) %>% 
  kable()
```

```{r}
df_raw %>% 
  summarise(across(everything(), ~ sum(is.na(.)))) %>% 
  pivot_longer(everything(), names_to = "Variable", values_to = "Missing") %>% 
  filter(Missing > 0)


df_raw %>% 
  filter(`What is your profession?` == "male")
```

**Missing values:** The `name` column has 1 missing entry, and 1 row had
`profession = "male"` (a clear data-entry error, recoded to `NA`). All
other key variables are complete (0 missing values). Hour variables are
encoded as intervals (e.g. `"3-4"` hours) rather than continuous values,
which limits precision.

## Distribution plot

```{r fig-distribution}
#| label: fig-dist
#| fig-cap: "Distribution of daily social media usage and age by self-reported addiction status"
#| fig-width: 10
#| fig-height: 4

p1 <- df %>% 
  filter(!is.na(addicted)) %>% 
  mutate(
    social_hours = factor(social_hours, levels = c("1-2","3-4","4-5","5-6","6+"))
  ) %>% 
  count(social_hours, addicted) %>% 
  group_by(social_hours) %>% 
  mutate(pct = n / sum(n)) %>% 
  ggplot(aes(x = social_hours, y = pct, fill = addicted)) +
  geom_col(position = "fill") +
  scale_y_continuous(labels = percent_format()) +
  scale_fill_manual(values = c("yes" = "#E07070", "no" = "#70B8E0", "not sure" = "#C0C0C0")) +
  labs(
    title = "Addiction by Daily Screen Time",
    x = "Hours/day on social media",
    y = "Proportion",
    fill = "Addicted?"
  ) +
  theme_minimal()

p2 <- df %>% 
  ggplot(aes(x = age)) +
  geom_histogram(binwidth = 2, fill = "#4C9BE8", colour = "white") +
  labs(title = "Age Distribution of Respondents", x = "Age (years)", y = "Count") +
  theme_minimal()

p2 + p1
```

```{r}
#| fig-width: 14
#| fig-height: 16
p1 <- ggplot(df, aes(
    x = social_hours,
    fill = addicted
  )) +
  geom_bar(position = "dodge") +
  theme_minimal() +
  labs(
    title = "Relationship: Social Media Usage vs. Self-Reported Addiction",
    x = "Average Daily Hours Spent on Social Media",
    y = "Number of Respondents",
    fill = "Considered Addicted?"
  )

p2 <- df %>% 
  filter(!is.na(social_hours), !is.na(family_hours)) %>% 
  mutate(
    social_hours = factor(social_hours, levels = c("1-2", "3-4", "4-5", "5-6", "6+")),
    family_hours = factor(family_hours, levels = c("1-2", "3-4", "4-5", "5-6", "6+"))
  ) %>% 
  ggplot(aes(x = social_hours, fill = family_hours)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = percent_format()) +
  labs(
    title = "Social Media Hours vs. Family Time",
    x = "Daily social media hours",
    y = "Proportion",
    fill = "Family hours/day"
  ) +
  theme_minimal()

p3 <- df %>%
  count(
    gender = gender,
    advertising_influence = ad_influence
  ) %>%
  group_by(gender) %>%
  mutate(percent = n / sum(n) * 100) %>%
  ggplot(aes(
    x = advertising_influence,
    y = percent,
    fill = gender
  )) +
  geom_col(position = "dodge") +
  labs(
    title = "Influence of Social Media Advertising by Gender",
    x = "Advertising influence on purchase decision",
    y = "Percentage within gender group",
    fill = "Gender"
  ) +
  theme_minimal()

p4 <- df %>% 
  mutate(
    age_group = cut(age, breaks = c(13, 20, 25, 30, 50),
                    labels = c("13-20", "21-25", "26-30", "30+")),
    social_hours = factor(social_hours, levels = c("1-2", "3-4", "4-5", "5-6", "6+"))
  ) %>% 
  ggplot(aes(x = age_group, fill = social_hours)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = percent_format()) +
  labs(
    title = "Age vs Daily Social Media Usage",
    x = "Age group",
    y = "Proportion",
    fill = "Hours/day"
  ) +
  theme_minimal()

p5 <- df %>% 
  filter(!is.na(addicted_bin)) %>% 
  ggplot(aes(x = social_hours_n, fill = factor(addicted_bin))) +
  geom_histogram(position = "fill", binwidth = 1) +
  scale_y_continuous(labels = percent_format()) +
  scale_fill_manual(values = c("0" = "#70B8E0", "1" = "#E07070"),
                    labels = c("0" = "No", "1" = "Yes")) +
  labs(
    title = "Proportion Addicted by Screen Time",
    x = "Hours/day (midpoint)",
    y = "Proportion",
    fill = "Addicted?"
  ) +
  theme_minimal()

(p4 + p1) / (p2 + p3) / p5
```

## Data quality

- Intervals, not continuous values: all time variables (`social_hours`,
  `family_hours`, `prof_hours`) are in the form of strings (for example,
  `“3-4”`).
- 78% of respondents are students, which limits generalisability to
  employed populations.
- One row has `profession = "male"` recoded to `NA`.
- The response options are Yes / No / Not sure. The "Not sure" group
  (`n = 125`, 24%) is substantial and will be excluded from binary
  logistic regression, reducing the effective sample.

# Analysis plan


## First research question

*Do younger people spend more time on social media?* Variables: `age`
and `social_hours`

Approach: examine whether younger users spend more time on social media.
We will use Exploratory Data Analysis (EDA). Since our variables are
largely categorical or ordinal, we will focus on frequency distribution
analysis and cross-tabulation. These methods are appropriate because
they allow for the identification of patterns and trends in user
behavior without requiring complex predictive modeling, which may be
over-fitted given the sample size of 522.\

Cleaning needed: convert `social_hours` to numeric, we will map the
string-based time intervals (e.g., "1-2", "6+") to numeric mid-points
(e.g., 1.5, 6.5) to enable basic arithmetic operations.

Risks: age range is 13–50 but highly concentrated at 22–27.

## Second research question

*Would people with higher screen time consider themselves addicted?*

Variables: `social_hours_n`, `benefits` and `addicted`

Approach: assess whether higher screen time is associated with feeling
addicted. Explore whether greater perceived benefits are linked to
higher addiction perception. We will use Pearson or Spearman correlation
(depending on normality).

Cleaning needed: parse `benefits` into a count variable. We will
normalize the categorical responses to ensure consistent formatting for
grouping and aggregation. Separating the used platforms and benefits in
individual columns in order for the analysis.

Risks: self-reported bias (screen time and addiction are subjective).
The source data was collected via a Google Form without strict
validation, leading to potential input errors. 

## Third research question

*Do people with higher social media screen time spend less time with their family?*

Variables: `social_hours_n`, `family_hours_n` and `profession`

Approach: scatter plot. Determine whether increased screen time is
associated with reduced family interaction. We will use Pearson or
Spearman correlation (depending on normality).

Risks: self-reported bias (screen time and addiction are subjective).\
Confounding variables: work/study requirements may affect screen time.\
Measurement inconsistency: different interpretations of "family time" or
"screen time". 

## Fourth research question

*Does the perceived influence of social media advertising on purchasing decisions differ between men and women?*

Variables: `ad_influence` and `gender`.

Approach: The analysis will use Exploratory Data analysis (EDA). 
Frequency tables and cross-tabulations will be used to examine how the advertising influence variable is distributed across genders. The results will then be visualized using group bar charts showing the percentage of each genders response. 
A Chi-square test of independence could also be conducted to assess if the differences between genders are statically significant. (Are they large enough to be significant at the 5% level?)


Cleaning transformation: The gender must be standardized, to ensure consistent coding of “Male” and “Female”.
The responses also have to be converted into an ordered factor so that categories are displayed from “strongly agree” to “Strongly disagree”. 

Risks: The data is self-reported so there is a bias that affects the responses. 
The number of male and female participants are uneven, which may influence comparisons and other results. 
There may be other factors like age, profession, time spent on media, etc, that could affect purchasing behavior. Gender alone may not explain the observed differences.


## Fifth research question

*Can we predict that someone considers themselves addicted to social media?*

Variables: outcome: `addicted_bin` (1 = Yes, 0 = No, "Not sure"
excluded) and predictors: `age`, `gender`, `social_hours_n`,
`forget_time`

Approach: logistic regression. Using logistic regression (if addiction
is binary) with predictors: screen time, age, gender, perceived
benefits, family time.

Cleaning needed: encode `forget_time` (Yes / Sometimes / No) and exclude
"Not sure" `addiction` rows. We will explicitly filter out or categorize
invalid entries in the What is your profession? column (e.g., the entry
"male") to maintain analytical integrity. We will normalize the
categorical responses to ensure consistent formatting for grouping and
aggregation.

Risks: self-reported bias (screen time and addiction are subjective).\
Confounding variables: since this is observational data, other variables
(e.g., age, academic workload, employment pressure) may influence the
results, making it difficult to establish direct causality between
social media usage and perceived addiction. The dataset is heavily
skewed toward students (\~78%). This limits the generalizability of
findings to other professional groups. The source data was collected via
a Google Form without strict validation, leading to potential input
errors.




# Workflow Organization

## Code Style

Style Guide: we will follow the tidyverse style guide, as it is the industry standard for R data projects, ensuring high readability and consistency.

Enforcement:

Linting: we will use the styler package to automatically format code according to the style guide before each commit.

Automation:We will integrate a pre-commit hook that runs the linter, preventing any non-compliant code from entering the repository.                       

## Packages

Core packages: - `tidyverse` - `scales` - `patchwork` - `knitr` /
`kableExtra` - `broom` - `pROC`

We will use **`renv`** to lock package versions, ensuring the project
renders identically across machines. The `renv.lock` file will be
committed to the repo, large cache folders will be excluded via
`.gitignore`.

## Git Workflow
Branching Strategy: We will follow a Feature Branch Workflow.
The main branch will always contain stable, production-ready code.
Every new task (e.g., "Add data cleaning script," "Create correlation plot") will be developed on a specific branch (e.g., feature/data-cleaning).
Once a feature is complete and passes linting, we will perform a Pull Request to merge it into main. This allows for code review and history tracking.

## gitignore
To keep the repository lightweight and secure, the .gitignore file will exclude the following:

Raw Data: Large or sensitive source files (e.g., Social Media Impact on Daily Lives.csv) should generally be kept out of version control if they are large or private, or managed via Data Version Control (DVC) if they change frequently.

Environment Files: Any .env files containing API keys, database credentials, or secret configuration variables (to prevent hardcoding sensitive information).

Rendered Outputs: Temporary files, plots, or HTML reports generated by the analysis (e.g., .html, .pdf, .png).

Project Cache: Local folder structures like .Rproj.user/ or the .Rhistory file.
