---
title: "Workflow Organization"
---

```{r setup}
#| include: false
source("setup.R")
```
# Workflow Organization

## Code Style

Style Guide: we will follow the tidyverse style guide, as it is the industry standard for R data projects, ensuring high readability and consistency.

Enforcement:
  
  Linting: we will use the styler package to automatically format code according to the style guide before each commit.

Automation:We will integrate a pre-commit hook that runs the linter, preventing any non-compliant code from entering the repository.                       

## Packages

Core packages: - `tidyverse` - `scales` - `patchwork` - `knitr` /
  `kableExtra` - `broom` - `pROC`

We will use **`renv`** to lock package versions, ensuring the project
renders identically across machines. The `renv.lock` file will be
committed to the repo, large cache folders will be excluded via
`.gitignore`.

## Git Workflow
Branching Strategy: We will follow a Feature Branch Workflow.
The main branch will always contain stable, production-ready code.
Every new task (e.g., "Add data cleaning script," "Create correlation plot") will be developed on a specific branch (e.g., feature/data-cleaning).
Once a feature is complete and passes linting, we will perform a Pull Request to merge it into main. This allows for code review and history tracking.

## gitignore
To keep the repository lightweight and secure, the .gitignore file will exclude the following:
  
  Raw Data: Large or sensitive source files (e.g., Social Media Impact on Daily Lives.csv) should generally be kept out of version control if they are large or private, or managed via Data Version Control (DVC) if they change frequently.

Environment Files: Any .env files containing API keys, database credentials, or secret configuration variables (to prevent hardcoding sensitive information).

Rendered Outputs: Temporary files, plots, or HTML reports generated by the analysis (e.g., .html, .pdf, .png).

Project Cache: Local folder structures like .Rproj.user/ or the .Rhistory file.

## Write and sanity-check a model specification

## Question
**Can we predict that someone considers themselves addicted to social media?**\

### Model Specification

* **Response Variable:** `addicted_bin` (Binary: 1 = Yes, 0 = No). *Note: "Not sure" responses are treated as NA and excluded from this binary model.*
  * **Predictor(s):** `social_hours_n` (Numeric), `age` (Numeric), `gender` (Categorical), and `forget_time` (Categorical).
* **Method:** Multiple Logistic Regression.
* **Expected Output:** We expect higher values in `social_hours_n` and a "Yes" in `forget_time` to be positively correlated with the probability of considering oneself addicted (Log-Odds > 0).
* **Interpretation:** The coefficients will show us how strongly each predictor influences the likelihood of perceived addiction, holding all other variables constant.

### Sanity Check (Data Validation)
```{r}
#| label: sanity-check-addiction-model

# 1. Do the variables exist, with the right type?
glimpse(df)

# 2. Is there enough variation and coverage to estimate what we want?
df |> count(addicted_bin)          
df |> count(gender)
df |> count(forget_time)

# 3. Missingness check for numeric variables
df |> summarise(across(where(is.numeric), \(x) sum(is.na(x))))
```
